---
title: "CCF_and_COR_for_ARMAX"
author: "Sky Yun"
date: "2026-04-07"
output: pdf_document
---

# Reading in Dataset
```{r}
source("data_cleaning_and_dependencies.R")
```

## CCF and Correlation Plot
**Colinearity Between LFP and WP**
```{r}
# Collinearity between LFP and WP
# Pre-Whitening Stage

corr_data <- select(train_df, -observation_date)
diff_data <- data.frame(
  diff_UR = diff(data[, UR_colname]),
  diff_LFP = diff(data[, LFP_colname]),
  diff_WAP = diff(data[, WAP_colname]),
  diff_IR = diff(data[, IR_colname])
)

colnames(corr_data) <- c("UR", "LFP", "WAP", "IR")
corr_table <- cor(diff_data)
corrplot(corr_table, method='number')

corr_table
```
- We decided to exclude working_pop_total from our regression model due to collinearity between unemployment_rate and labor_force_total. 
- Hence, we only examine interest rate and labor force participation in our ARMAX model. 

**CCF Plots for each of the predictors**
```{r}
# Check if x_t leads y_t (y_t, x_t (Labor Force Participation)) 
# Due to the non-stationarity of y_t, differencing helps avoid spurious correlations
ccf(diff(train_df$interest_rate), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
ccf(diff(train_df$labor_force_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment rate) and diff(Labor Force Participation) CCF")
ccf(diff(train_df$working_pop_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
```
For all CCF values, we assume that the predictors will lead unemployment rate so we only observe lagged $h < 0$ values