---
title: "ARIMA and ARMAX"
author: "Sky Yun"
date: "2026-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data Cleaning
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggcorrplot))

unempl <- read.csv("data/unemployment.csv", header=TRUE)
LF_total <- read.csv("data/LFPR_total.csv", header=TRUE)
WP_total <- read.csv("data/WP_total.csv", header=TRUE)
int_rate <- read.csv("data/interest_rate.csv", header=TRUE)


colnames(unempl) <- c("observation_date", "unemployment_rate")
colnames(LF_total) <- c("observation_date", "labor_force_total")
colnames(WP_total) <- c("observation_date", "working_pop_total")
colnames(int_rate) <- c("observation_date", "interest_rate")

# Combine dataframes into on
data <- merge(unempl, LF_total, id="observation_date") %>% 
        merge(WP_total, id="observation_date") %>%
        merge(int_rate, id="observation_date")

# Create time Series Objects
start_date <- c(1976, 1)
train_end_date <- c(2015, 12)
holdout_start_date <- c(2016, 1)

unempl_ts <- ts(data$unemployment_rate, start=start_date, frequency=12)
train_ts <- window(unempl_ts, end=c(2015, 12))
holdout_ts <- window(unempl_ts, start=c(2016, 1))

n_train = length(train_ts)
n = nrow(data)

train_df <- data[1:n_train, ]
holdout_df <- data[n_train+1:n, ]
```

# Exploratory Analysis

**Colinearity Between LFP and WP**
```{r}
# Collinearity between LFP and WP
corr_table <- cor(select(train_df, -observation_date))
```
- We decided to exclude working_pop_total from our regression model due to collinearity between unemployment_rate and labor_force_total. 
- Hence, we only examine interest rate and labor force participation in our ARMAX model. 
```{r}
# Check if x_t leads y_t (y_t, x_t (Labor Force Participation)) 
# Due to the non-stationarity of y_t, differencing helps avoid spurious correlations
ccf(diff(train_df$interest_rate), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate Cross-Correlation)")
ccf(diff(train_df$labor_force_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment rate) and diff(Labor Force Participation Cross-correlation)")
ccf(diff(train_df$working_pop_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate Cross-Correlation)")

# WP_total
```
For all CCF values, we assume that the predictors will lead unemployment rate so we only observe lagged $h < 0$ values

# Check

# ARIMA
```{r}
arima_fit <-  auto.arima(train_ts, stationary=FALSE, seasonal=TRUE)
holdout_pred <- predict(arima_fit, n.ahead=length(test_ts))
```



## Training:
```{r}
# ARMAX
# resid <- arima_fit$residuals
# resid_arima <- auto.arima(resid, seasonal=FALSE)
# 
# summary(resid_arima)
# 


LF_ts <- ts(data$labor_force_total, start=start_date, frequency=12)
train_LF_ts <- window(LF_ts, end=train_end_date)
holdout_LF_ts <- window(LF_ts, start=holdout_start_date)

predictors_train = data[1:n_train,c("labor_force_total", "interest_rate")]

armax <- auto.arima(train_ts, xreg=as.matrix(predictors_train))
```




```{r}
lm_model <- lm(unemployment_rate ~ labor_force_total + interest_rate, data=train_df)
resid = lm_model$residual

plot(resid)
auto.arima(resid)
```

## Testing: Evaluating RMSE Values
```{r}
predictors_holdout = data[n_train+1:n, c("labor_force_total", "interest_rate")]
y_armax_pred <- as.data.frame(f)[["Point Forecast"]]
```




