---
title: "ARMAX"
output: pdf_document
---

# Data and Dependencies
```{r}
source("data_cleaning_and_dependencies.R")
```

## CCF and Correlation Plot
**Colinearity Between LFP and WP**
```{r}
# Collinearity between LFP and WP
# Pre-Whitening Stage

corr_data <- select(train_df, -obs_date_colname)
diff_data <- data.frame(
  diff_UR = diff(data[, UR_colname]),
  diff_LFP = diff(data[, LFP_colname]),
  diff_WAP = diff(data[, WAP_colname]),
  diff_IR = diff(data[, IR_colname])
)

colnames(corr_data) <- c("UR", "LFP", "WAP", "IR")
corr_table <- cor(diff_data)
corrplot(corr_table, method='number')
corr_table
```

**CCF Plots for each of the predictors**
```{r}
# Check if x_t leads y_t (y_t, x_t (Labor Force Participation)) 
# Due to the non-stationarity of y_t, differencing helps avoid spurious correlations
ccf(diff(train_df$int_rates), diff(train_df$unemployment), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
ccf(diff(train_df$labor_part), diff(train_df$unemployment), 
    main="diff(Unemployment rate) and diff(Labor Force Participation) CCF")
ccf(diff(train_df$work_pop), diff(train_df$unemployment), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
```
For all CCF values, we assume that the predictors will lead unemployment rate so we only observe lagged $h < 0$ values


# ARIMAX Model Fitting
```{r}
# build matrix of explanatory variables for ARIMAX
xreg_all <- as.matrix(data[, c("labor_part", "int_rates")])
xreg_train <- xreg_all[1:n_train, , drop = FALSE]
xreg_holdout <- xreg_all[(n_train + 1):n, , drop = FALSE]
```

**ARIMAX**
```{r}
arimax_auto <- auto.arima(
  train_ts,
  xreg = xreg_train,
  seasonal = TRUE
)

#auto.arimax chose ARIMA(2,0,2)(1,0,0)[12] errors 
summary(arimax_auto) # 3.723267
# (1,0,2)(1,0,0)12 2.754148.    -255.06
# (2,0,1)(1,0,0)12 3.767669.    -303.66
# (2,1,2)(1,1,0)12 3.546538.    -37.06
# (3,0,2)(1,0,0)12 3.536629.    -304.3
# (2,0,2)(2,0,1)12 3.332236.    -249.57
# (2,0,2)(2,0,0)12 3.706091.    -317.6
```
**CSS Method**
```{r}
# arimax CSS method
arimax_css <- Arima(
  train_ts,
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  include.mean = TRUE,
  method = "CSS",
  xreg = xreg_train
)
print(arimax_css)
```
**ML Method**
```{r}
# arimax ML method
arimax_ml <- Arima(
  train_ts,
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  include.mean = TRUE,
  xreg = xreg_train
)
print(arimax_ml)
```

# Forecasting
## ARIMAX 1-step ahead forecast FUnction for holdout set
```{r}
arimax_fc <- function(tsdata, ntrain, order, seasonal, method, traincoef,
                      include.mean, xreg, iprint = FALSE) {
  obj <- arima(
    tsdata,
    order = order,
    seasonal = seasonal,
    init = traincoef,
    fixed = traincoef,
    method = method,
    include.mean = include.mean,
    xreg = xreg,
    transform.pars = FALSE,
    optim.control = list(maxit = 0)
  )
  
  fc <- tsdata - obj$residuals
  ntotal <- length(tsdata)
  holdout_fc <- fc[(ntrain + 1):ntotal]
  holdout <- tsdata[(ntrain + 1):ntotal]
  
  if (iprint) {
    print(cbind(holdout, holdout_fc))
  }
  
  rmse <- sqrt(mean((holdout - holdout_fc)^2))
  list(rmse = rmse, fc = holdout_fc)
}
```



**Forecast for CSS ARIMAX**
```{r}
# h-step forecast css
fc_css <- forecast(arimax_css, 
                   h = length(holdout_ts), 
                   xreg = xreg_holdout)
rmse_css <- sqrt(mean((holdout_ts - fc_css$mean)^2))
rmse_css # 3.561992

# 1-step forecast css
css_fc <- arimax_fc(
  tsdata = unempl_ts,
  ntrain = n_train,
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "CSS",
  traincoef = arimax_css$coef,
  include.mean = TRUE,
  xreg = xreg_all
)
css_fc$rmse # 1.11026
```


**Forecast for ML ARIMAX**
```{r}
# h-step forecast ml
fc_ml <- forecast(arimax_ml, 
                          h = length(holdout_ts), 
                          xreg = xreg_holdout)
rmse_ml <- sqrt(mean((holdout_ts - fc_ml$mean)^2))
rmse_ml # 3.773121

# 1-step forecast ml
css_fc <- arimax_fc(
  tsdata = unempl_ts,
  ntrain = n_train,
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = arimax_ml$coef,
  include.mean = TRUE,
  xreg = xreg_all
)
css_fc$rmse # 1.105931
```

## Forecast for ARIMAX Models with Perturbed Values

```{r}
# 2. change p, P, q, Q values to find model that lowers holdout RMSE
arimax_changed <- Arima(
  train_ts,
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  xreg = xreg_train
)
summary(arimax_changed)

fc_changed <- forecast(arimax_changed, 
                h = length(holdout_ts), 
                xreg = xreg_holdout)
rmse_changed <- sqrt(mean((holdout_ts - fc_changed$mean)^2))
rmse_changed # 2.754148
```


**arimax CSS method for changed model**
```{r}
arimax_changed_css <- Arima(
  train_ts,
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "CSS",
  include.mean = TRUE,
  xreg = xreg_train
)
print(arimax_changed_css)

# h-step changed css forecast
fc_changed_css <- forecast(arimax_changed_css, 
                   h = length(holdout_ts), 
                   xreg = xreg_holdout)
rmse_changed_css <- sqrt(mean((holdout_ts - fc_changed_css$mean)^2))
rmse_changed_css # 2.75788

# 1-step changed css forecast
onestep_fc_changed_css <- arimax_fc(
  tsdata = unempl_ts,
  ntrain = n_train,
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = arimax_changed_css$coef,
  include.mean = TRUE,
  xreg = xreg_all
)
onestep_fc_changed_css$rmse # 0.9944612
```
**arimax ML method for changed model**
```{r}
arimax_changed_ml <- Arima(
  train_ts,
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  include.mean = TRUE,
  xreg = xreg_train
)
print(arimax_changed_ml)

# h-step changed ml forecast
fc_changed_ml <- forecast(arimax_changed_ml, 
                   h = length(holdout_ts), 
                   xreg = xreg_holdout)
rmse_changed_ml <- sqrt(mean((holdout_ts - fc_changed_ml$mean)^2))
rmse_changed_ml # 2.754148

# 1-step changed ml forecast
onestep_fc_changed_ml <- arimax_fc(
  tsdata = unempl_ts,
  ntrain = n_train,
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = arimax_changed_ml$coef,
  include.mean = TRUE,
  xreg = xreg_all
)
onestep_fc_changed_ml$rmse # 0.9944082
```




