---
title: "CCF_and_COR_for_ARMAX"
author: "Sky Yun"
date: "2026-04-07"
output: pdf_document
---

# Data Cleaning
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggcorrplot))

#
unempl <- read.csv("data/unemployment.csv", header=TRUE)
LF_total <- read.csv("data/LFPR_total.csv", header=TRUE)
WP_total <- read.csv("data/WP_total.csv", header=TRUE)
int_rate <- read.csv("data/interest_rate.csv", header=TRUE)

# Renaming col-names
colnames(unempl) <- c("observation_date", "unemployment_rate")
colnames(LF_total) <- c("observation_date", "labor_force_total")
colnames(WP_total) <- c("observation_date", "working_pop_total")
colnames(int_rate) <- c("observation_date", "interest_rate")

# Combine dataframes into on
data <- merge(unempl, LF_total, id="observation_date") %>% 
        merge(WP_total, id="observation_date") %>%
        merge(int_rate, id="observation_date")

# Create time Series Objects
start_date <- c(1976, 1)
train_end_date <- c(2015, 12)
holdout_start_date <- c(2016, 1)

unempl_ts <- ts(data$unemployment_rate, start=start_date, frequency=12)
train_ts <- window(unempl_ts, end=c(2015, 12))
holdout_ts <- window(unempl_ts, start=c(2016, 1))

n_train = length(train_ts)
n = nrow(data)

# Creating dataframe Object
train_df <- data[1:n_train, ]
holdout_df <- data[n_train+1:n, ]

head(train_df)
```

## CCF and Correlation Plot
**Colinearity Between LFP and WP**
```{r}
# Collinearity between LFP and WP
corr_table <- cor(select(train_df, -observation_date))corr_table
```
- We decided to exclude working_pop_total from our regression model due to collinearity between unemployment_rate and labor_force_total. 
- Hence, we only examine interest rate and labor force participation in our ARMAX model. 

**CCF Plots for each of the predictors**
```{r}
# Check if x_t leads y_t (y_t, x_t (Labor Force Participation)) 
# Due to the non-stationarity of y_t, differencing helps avoid spurious correlations

ccf(diff(train_df$interest_rate), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
ccf(diff(train_df$labor_force_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment rate) and diff(Labor Force Participation) CCF")
ccf(diff(train_df$working_pop_total), diff(train_df$unemployment_rate), 
    main="diff(Unemployment Rate) and diff(Interest_rate) CCF")
```
For all CCF values, we assume that the predictors will lead unemployment rate so we only observe lagged $h < 0$ values